Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory

### 2021-01-15 papers/thesis-kirstie/thesis.tex
edits

A phase transition occurs along a phase boundary line on a phase diagram
such as can be seen in the pressure-temperature (P-T) phase diagram
shown in Figure~\ref{fig:P-T_Diagram}. 
%
%Points along the phase boundary that separates the solid and the fluid phases 
%indicate a phase transition where the solid and fluid phases coexist, 
%and show that both the solid and the fluid have the same temperature (necessary 
%for thermal equilibrium)
%and the same pressure (necessary for mechanical 
%equilibrium)at a phase transition, and the same chemical potential 
%(necessary for diffusive equilibrium). 
%
This phase diagram is 
representative of some substance 
which exhibits both attraction and repulsion. However, since 
the atoms in a WCA fluid do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, there will simply
be a fluid.
The phase diagram indicates the equilibrium state a substance will be in at
different temperatures and pressures.
At a phase transition, both the fluid and the solid have 
the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium) and the same chemical potential (necessary for diffusive equilibrium).
%The temperature , pressure, and chemical potential will be the same for both
%the fluid and the solid at a fluid-solid phase transition.

@@

        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  

@@

      w_3 &= \Theta(r_{ij}-R)  \label{scalar-weight3} \\

@@

     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}                                                              

@@

@@

Lifting the restriction to the low density limit, Rosenfeld derived an
expression for the excess Helmholtz free energy by noting that the higher
order terms in the virial expansion can all be written in terms of an 
``exclusion volume" (given by Equation~\ref{exclusion-volume}), 
and an ``overlap volume" which is the volume of intersection between two 
overlapping spherical regions.
The exclusion volume is the spherical volume about the center of one sphere 
which the center of another sphere can not enter, and is used to describe
terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. Used in conjunction 
with the overlap volume, the higher order terms in the expansion
in Equation~\ref{eq:Fexcess-simplified}~\cite{rosenfeld1988} can be described. 
%
The resulting expression for the excess Helmholtz free energy with Tarazona's $\Phi_3$
(see Introduction) is given by
\begin{align}
    F_{excess}[n(\vec{r})] &= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r} \\
    %\Phi_1 &= -n_{0}ln(1-n_{3})  \\
    %\Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
    %\Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}
    \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
    \Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
    \Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align}
%where in addition to the weight functions already defined,  
where the weighted densities are given by Equation~\ref{weighted_densities} with 
weight functions
\begin{align} 
      w_3 &= \Theta(r-R)  \label{scalar-weight3} \\
      w_2 &= \delta(r-R)  \\
      w_1 &= \frac{w_2}{4\pi r}  \\
      w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}  
\end{align}
\begin{align}      
      \vec {w}_2 &= w_2~\hat r \\
      \vec {w}_1 &= w_1~\hat r \\
      \overleftrightarrow{w}_{m2}(\vec{r}) &= w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
\end{align}
%and a tensor weight function given by
%\begin{align}\label{eq:tensorweight-methods}
     %\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
%\end{align}
%The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
%all depend on the spatially varying number density profile $n(\vec r)$.
%\begin{align}%\label{eq:numdenprofile}
	%n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(\vec{r}-\vec{r'})d{\vec{r'}}}
%\end{align}
This result reduces to the same result obtained from Scaled Particle Theory 
for a homogeneous hard-sphere fluid.

>>>>

A phase transition occurs along a phase boundary line %on a phase diagram
such as can be seen in the pressure-temperature (P-T) phase diagram
shown in Figure~\ref{fig:P-T_Diagram}. 
%
%Points along the phase boundary that separates the solid and the fluid phases 
%indicate a phase transition where the solid and fluid phases coexist, 
%and show that both the solid and the fluid have the same temperature (necessary 
%for thermal equilibrium)
%and the same pressure (necessary for mechanical 
%equilibrium)at a phase transition, and the same chemical potential 
%(necessary for diffusive equilibrium). 
%
This phase diagram is 
representative of some substance 
which exhibits both attraction and repulsion. However, since 
the atoms in a WCA fluid do not exhibit attraction there will not be a 
dividing line between the gas and liquid phases, there will simply
be a fluid.
The phase diagram indicates the equilibrium state a substance will be in at
different temperatures and pressures.
At a phase transition, both the fluid and the solid have 
the same temperature (necessary 
for thermal equilibrium), the same pressure (necessary for mechanical 
equilibrium) and the same chemical potential (necessary for diffusive equilibrium).
%The temperature , pressure, and chemical potential will be the same for both
%the fluid and the solid at a fluid-solid phase transition.

@@

        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  \\

@@

      w_3 &= \Theta(r_{ij}-R)  \label{scalar-weight3} 

@@

     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}         \\ 

@@

                %&\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
                %&\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)\left(\int w_0(\vec u - \vec r_i)w_3(\vec u - \vec r_j) d\vec u \color{white}\right)  \nonumber \\
                %&+ \color{white}\left(\color{black}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec u ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)  \\  

@@

Lifting the restriction to the low density limit, Rosenfeld derived an
expression for the excess Helmholtz free energy by noting that the higher
order terms in the virial expansion can all be written in terms of an 
``exclusion volume" (given by Equation~\ref{exclusion-volume}), 
and an ``overlap volume" which is the volume of intersection between two 
overlapping spherical regions.
The exclusion volume is the spherical volume about the center of one sphere 
which the center of another sphere can not enter, and is used to describe
terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. Used in conjunction 
with the overlap volume, the higher order terms in the expansion
in Equation~\ref{eq:Fexcess-simplified}~\cite{rosenfeld1988} can be described. 
%
The resulting expression for the excess Helmholtz free energy with Tarazona's $\Phi_3$
(see Introduction) is given by
\begin{align}
    F_{excess}[n(\vec{r})] &= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r} \\
    %\Phi_1 &= -n_{0}ln(1-n_{3})  \\
    %\Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
    %\Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}
    \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
    \Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
    \Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align}
%where in addition to the weight functions already defined,  
where the weighted densities are given by Equation~\ref{weighted_densities} with 
weight functions
\begin{align} 
      w_3 &= \Theta(r-R)  \label{scalar-weight3} \\
      w_2 &= \delta(r-R)  \\
      w_1 &= \frac{w_2}{4\pi r}  \\
      w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}  \\
%\end{align}
%\begin{align}      
      \vec {w}_2 &= w_2~\hat r \\
      \vec {w}_1 &= w_1~\hat r \\
      \overleftrightarrow{w}_{m2}(\vec{r}) &= w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
\end{align}
%and a tensor weight function given by
%\begin{align}\label{eq:tensorweight-methods}
     %\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
%\end{align}
%The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
%all depend on the spatially varying number density profile $n(\vec r)$.
%\begin{align}%\label{eq:numdenprofile}
	%n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(\vec{r}-\vec{r'})d{\vec{r'}}}
%\end{align}
This result reduces to the same result obtained from Scaled Particle Theory 
for a homogeneous hard-sphere fluid.
